---
title: "TE Distribution and Normalizaiton (by class)"
author: "L.S."
date: "2021/3/3"
output: html_document:
    theme: cerulean
    toc: no
    number_sections: no
  pdf_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Setup
```{r}
rm(list = ls())
library(dplyr)
library(reshape2)
options(scipen = 999)
```

## Data prep.
### load the matrix
TE class were sorted by number, and TFs were sorted by looping z-score.
```{r}
load("./nromedpkn_class_matrix_sorted.Rdata")

## transform to long data format
dat <- melt(normpkn_cls_mx)
colnames(dat) <- c("class", "tf", "norm")
```
### Load looping z-score
```{r}
lp_zsc <- read.csv("../.././../../hs_getloops/intrachr/3.plot/exless_h.all.tf_des-meanz.v3.csv", header = T)

## group
lp_zsc$group <- ifelse(lp_zsc$allmean_z >= 0.8, "high",
                       ifelse(lp_zsc$allmean_z <= -0.8, "low",
                              "mid"))

## reverse the tf order for later heatmap usage
dat$tf <- factor(dat$tf,
                 levels = rev(colnames(normpkn_cls_mx)))

## assign colors to TFs in diff. groups
labelcol <- ifelse(lp_zsc$group == "high", "#CD5C5C",   #'high' = 'IndianRed'
                  ifelse(lp_zsc$group == "mid", "#DAA520", #'mid' = 'Goldenrod'
                         "#4682B4"))  #'low' = 'SteelBlue'
```


## Default matrix
### Barchart 
Stacked and percent stacked barplot to show the number and proportion.
Setup and color:
```{r}
library(ggplot2)

theme_set(
  theme_minimal() +
    theme(legend.position = "right",
          legend.key.size = unit(12, "pt"))
)

library(ggsci)
simpsons <- pal_simpsons("springfield")(16)
```
#### Stacked barchart
```{r}
# stack
p <- ggplot(data = dat,
            aes(x = tf, y = norm, fill = class))
p + geom_bar(stat = "identity", position = "stack") +
  labs(title = "TE Class Distribution",
       x = NULL, y = "Normed TE Number") +
  coord_flip() +
  scale_fill_manual(values = simpsons) +
  guides(fill = guide_legend(title = "TE Class", reverse = F)) +
  theme(axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 7, colour = rev(labelcol), face = "bold"),
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 6))
#ggsave(filename = "../../figure/norm_resize/tf_class_bar_stack.png", width = 7,height = 12, dpi = 1000)
#ggsave(filename = "../../figure/norm_resize/tf_class_bar_stack.pdf", width = 7,height = 12, dpi = 1000)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
